---
title: "PA1_template.Rmd PA1_template.Rmd"
output: md_document
---
# Reproducible Research Project 1
```{r message=FALSE}

library(plyr)
library(dplyr)
library(lattice)
library(ggplot2)
```

## Loading and Reading Data

```{r}

if(!file.exists("./data")){dir.create("./data")}
fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileurl, destfile = "./data/activity.zip")
unzip(zipfile = "./data/activity.zip", exdir = "./data")
        
activity <-read.csv("./data/activity.csv")

```
## Task 1: What is mean total number of steps taken per day?


**Calculate the total number of steps taken per day**

```{r echo=TRUE}

tot_steps <- aggregate(steps~date, activity, sum, na.action = na.pass)
head(tot_steps)
```
        

**Make a histogram of the total number of steps taken each day**

```{r echo=TRUE, warning = FALSE}

ggplot(tot_steps, aes(x = steps)) +
        geom_histogram(fill = "RED", binwidth = 1000) +
        labs(title = "Total number of Steps taken each day", x = "Steps per day", y = "Frequency")

```
        
**Calculate and report the mean and median of the total number of steps taken per day**

```{r echo=TRUE}

 mean_tot_steps <- mean(tot_steps$steps, na.rm = TRUE)
        median_tot_steps <- median(tot_steps$steps, na.rm = TRUE)     
        cat("Mean: ", mean_tot_steps, "Median: ", median_tot_steps)
```
        

##  Task 2: What is the average daily activity pattern?

**Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**

```{r}

### Aggregating the mean of Steps with Intervals and ignoring NAs
step_int <- aggregate(steps~interval, activity, mean, na.action = na.omit)

head(step_int)

ggplot(step_int, aes(x = interval , y = steps)) +
        geom_line(color="black", linewidth=1) + 
        labs(title = "Average Daily Steps per 5min Intervall", x = "Interval", y = "Avg. Daily Steps")

```

**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**

```{r}

max_inter <- subset(step_int, steps == max(steps), interval)
max_inter

```


## Task 3: Imputing missing values
        
**Calculate and report the total number of missing values in the dataset **

```{r}

sum(is.na(activity$steps))

```
        
**Devise a strategy for filling in all of the missing values in thedataset.** 
        
```{r}
#Strategy: Replace missing NA with mean of interval
        
        #Create new dataframe equal to initial activitiy_df, but replacing NA with mean of interval
        new_df <- activity
        no_NA <- subset(activity, !is.na(activity$steps))
        no_NA2 <- is.na(new_df$steps)
        avg_int <- tapply(no_NA$steps, no_NA$interval, mean, na.rm=TRUE, simplify=T)
        new_df$steps[no_NA2] <- avg_int[as.character(new_df$interval[no_NA2])]

```
        
**Create a new dataset that is equal to the original dataset but with the missing data filled in. **


```{r}

corrected_df <- aggregate(steps~date, new_df, sum, na.action = na.omit)

```
                
     
**Make a histogram of the total number of steps taken each day ** 

```{r}
ggplot(corrected_df, aes(x = steps)) +
        geom_histogram(fill = "red", binwidth = 750) +
        labs(title = "Total number of Steps taken each day", x = "Steps per day", y = "Frequency")
```


**Calculate and report the mean and median total number of steps taken  per day.**

```{r}
mean(corrected_df$steps)
median(corrected_df$steps)
```              



**Do these values differ from the estimates from the first part of the assignment?**

no difference

### Task 4: Are there differences in activity patterns between weekdays and weekends?

**Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.**. 
```{r}
new_df$date<-as.Date(new_df$date)

day_df<-new_df %>% 
        mutate(dayType= ifelse(weekdays(new_df$date)=="Saturday" | weekdays(new_df$date)=="Sunday", "Weekend", "Weekday"))

week_df <- aggregate(steps ~ dayType+interval, data=day_df, FUN=mean)
```


**Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken,   averaged across all weekday days or weekend days (y-axis)**

```{r}

 library(lattice)
                xyplot(steps ~ interval | factor(dayType),
                       layout = c(1, 2),
                       col = "red",
                       xlab="Interval",
                       ylab="Total steps",
                       type="l",
                       lty=1,
                       data=week_df)

```
